---
title: "Capstone Project II"
author: "Tarasov Oleksiy"
date: "6/2/2019"
output: html_document
---

# Introduction
This report is part of the HarvardX Certification of the edX HarvardX: PH125.9x Data Science: Capstone Online Training
The data was taken in acccordance with the proposal of the course from Kaggle curated list of datasets
link https://www.kaggle.com/annavictoria/ml-friendly-public-datasets?utm_medium=email&utm_source=intercom&utm_campaign=data+projects+onboarding
The following dataset were chosen for the analysis

[<Adult Census Income>](https://www.kaggle.com/uciml/adult-census-income) https://www.kaggle.com/uciml/adult-census-income
in case that connection to the Kaggle will not work  I have copied the file to my github directory 
https://github.com/alexej-tarasov/CapstoneII file adult.csv

# Overview
The target cell is the column which predicts whether the person has incomes more or less then 50000 USD annually


# Executive Summary


# Methods/analysis 
This section  explains the process and techniques used, such as data cleaning, data exploration and visualization, any insights gained, and your modeling approach;
First library for analysis should be activated
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(caret)
library(tidyr)
library(plotly)
library(tidyverse)
library(lubridate)
library(broom)
library(ggplot2)
```

The data will be read 
```{r}
data<-read.csv("~/CapstoneII/CapstoneII/adult.csv")

```
The data has following attributes

**Target income**: >50K, <=50K

**age**: continuous

**workclass**: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked

**fnlwgt**: continuous

**education**: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool

**education-num**: continuous

**marital-status**: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse

**occupation**: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces

**relationship**: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried

**race**: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black

**sex**: Female, Male

**capital-gain**: continuous

**capital-loss**: continuous

**hours-per-week**: continuous

**native-country**: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands

# Results section
# Conclusion section.
